---
title: 'data visualization #awesome'
author: "Jared Petry"
date: "2022-08-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse) 
library(here) 
library(janitor) 

#cool graph libraries from gg__
library(patchwork)
library(ggrepel)
library(gghighlight) 
library(paletteer)
library(ggExtra) 
library(ggbeeswarm) 
library(gapminder) 
library(sf)

```

# Graphs

Read in the data 
```{r}
lizards <- read_csv(here::here("data_tidy", "lizards.csv"))
```

make some graphs 
```{r}
ggplot(data = lizards) +geom_line(aes(x = total_length, y = weight))
```

create historgram 
```{r}
ggplot(data = lizards, aes(x = total_length)) +
  geom_histogram(color = "orange", 
                 fill = "purple", 
                 size = 0.3, 
                 linetype = "dotted")

    #there are rules for determining the optimal number of bins for a given sample size that we'll learn later
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) + 
  geom_point(shape = 22,       #shape of the dots 
             color = "cyan4",   #color of the dots' border
             fill = "yellow",   #color of the dot fill
             size = 4,      #size of the dots
             alpha = 0.4)  #transparency
```

map asthetics based on variables
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) + 
  geom_point(aes(color = common_name,    #--- color is mapped based on common name
                 size = total_length,    #size of the dot is based on the total length 
                 alpha = 0.4)) +
  theme_minimal() +          #many different themes to choose from but you can also create your own 
  facet_wrap(~common_name, scales = "free") #scales free means that the axis's on the faceted graphs will be differrent                                               based on the values in each 
```

find total counts of lizards using common name as grouping variable 
```{r}
lizards_totals <- lizards |> 
  group_by(common_name) |> 
  summarize(counts = n())

##--- to do the exact same thing 
#dplyr::count()
lizard_counts <- lizards |> 
  count(common_name)
```

Make a bar plot , column graph that plots this count thing we just created 

```{r}
ggplot(data = lizards_totals, aes(y = fct_reorder(common_name, counts), x = counts)) +
  geom_col()
```

but the plot makes the common names in alphabetical order, but we would rather have the order based on the count variable 

this is where factors come in handy 

#factors 

fct_reorder() within aes in gg plot .. specify (this variable, order based on this variable)
within the farecats package 

# more graphs 

create scatterplots by species for total length vs weight
color will differ based on species (common name) 
change theme 
make each faceted panel have flexible axis limits (scale is free)
and then change whatever else you want

```{r}
liz_length_weight <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) + 
  theme_get() + 
  facet_wrap(~common_name, scales = "free") + 
  labs(title = "Really sick lizard graph", subtitle = "a smaller title", caption = "data liscensing", 
       x= "lizard length", y= "lizard weight")

liz_length_weight
```

## Fun with scales 

```{r}
liz_length_weight_2 <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) + 
  scale_color_gradientn(colors = c("magenta", "green", "dodgerblue"))

liz_length_weight_2
```

```{r}
liz_length_weight_3 <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) + 
    scale_color_steps2(low = "green", 
                       mid = "yellow", 
                       high = "red", 
                       midpoint = 30, 
                       breaks = c(5, 15, 25))

liz_length_weight_3
```

```{r}
liz_length_weight_4 <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = total_length)) + 
  scale_color_stepsn(colors = c("orange", "red", "purple"), 
                     breaks = seq(from = 0, to = 300, by = 60))

liz_length_weight_4
```

horizontal boxplot y has common name and x has total_length with boxplot fill chnaging color based on the common name 
```{r}
lizards_median <- lizards |>  
  mutate(common_name = fct_reorder(common_name, total_length, .fun = mean))

lizard_box <- ggplot(data = lizards_median, 
                    aes(y = common_name, 
                        x = total_length)) + 
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) + 
  scale_fill_paletteer_d(palette = "yarrr::usualsuspects") + 
  theme_minimal()

lizard_box
```

### everything example
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  theme(panel.grid.major.x  = element_line(color = "red"), 
        panel.grid.major.y = element_line(color = "green"), 
        axis.title.x = element_text(color = "purple"), 
        axis.text.y = element_text(color = "orange"), 
        text = element_text(size = 18), 
        panel.background = element_rect(color = "yellow", fill = "pink")) + 
  annotate("text", x = 300, y = 50, label = "YEWWW", color = "blue")
```

### Repulsive labels
```{r}
wws_lizards <- lizards |> 
  filter(common_name == "western whiptail", site == "sand")

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  geom_text_repel(aes(label = toe_num)) #automatic repulsion from the points so you can read the values
```

use ggrepel to label the points on a scatterplot 
```{r}
gapminder |>  
  filter(year == 2002, continent == "Europe") |> 
  ggplot(aes(x = gdpPercap, y = lifeExp)) + 
  geom_point() + 
  geom_text_repel(aes(label = country), size = 3)
```
how to highlight and add a label to a point in the graph 
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() + 
  gghighlight(weight > 30, label_key = toe_num)
```

# Make a map 
fuck yea!

sf stands for simple features... way of storing spatial data and using it the same way you are using a dataframe
read in a KML file... will store polygon as a list of lat/long info that would be like a connect the dot for that polygon

```{r}
jornada_veg <- read_sf(here::here("data_raw", "spatial_vegetation", "doc.kml"))

ggplot() + 
  geom_sf(data = jornada_veg, aes(fill = Name), color = NA) + 
  theme_minimal() + 
  scale_fill_paletteer_d(palette = "ggthemes::manyeys")
```

