---
title: "make a heatmap"
author: "Jared Petry"
date: "2022-08-19"
output: html_document
---

# heatmap

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse) 
library(here) 
library(janitor) 

#cool graph libraries from gg__
library(patchwork)
library(ggrepel)
library(gghighlight) 
library(paletteer)
library(ggExtra) 
library(ggbeeswarm) 
library(gapminder) 
library(sf)

```

read in
```{r}

lizards <- read_csv(here::here("data_tidy", "lizards.csv"))
```

cleanup
```{r}
lizards_counts <- lizards |> 
  mutate(date = lubridate::mdy(date)) |> 
  count(year = lubridate::year(date), common_name) |> 
  drop_na()
```

graph: 
HEATMAP
```{r}
ggplot(data = lizards_counts, aes(x = year, y = common_name)) +
  geom_tile(aes(fill = n), show.legend = FALSE) +
  geom_text(aes(label = n), color = "white", size = 3) + 
  scale_fill_gradientn(colors = c("navy", "red", "orange")) + 
  theme_minimal()
```

# beeswarm plot 
i particularly like this one for when you have count data 
more accurate than a jitter plot but they only work well for relatively small sample sizes 


create a subset 
```{r}
whiptails <- lizards |> 
  filter(common_name == "western whiptail") |> 
  drop_na(total_length, weight)
```

graph it (can add a boxplot in there too if you want)

```{r}
whiptail_bee <- ggplot(data = whiptails, aes(x = sex, y = weight)) +
  geom_beeswarm(size = 0.5) + 
  geom_boxplot(fill = NA)
```

create scatter plot with weight on y axis and length on x axis with point color depending on sex 
```{r}
whiptail_plot <- ggplot(data = whiptails, aes(x = total_length, y = weight)) + 
  geom_point(aes(color = sex)) + 
  theme(legend.position = "bottom")

#---- put marginal plots on the edges of the graph
ggMarginal(whiptail_plot, type = "boxplot", groupColour = TRUE)
```
# patchwork
fully customize the output of compound figures with PATCHWORK
put figures together like PEMDAS

```{r}
#---- plus sign makes columns of things (puts them side by side)
(whiptail_bee + whiptail_plot)

#--- division symbol puts things in rows of things 
(whiptail_bee / whiptail_plot)
```

